---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r}
if (!file.exists("activity.csv")) {
  unzip("activity.zip")
}
original <- read.csv("activity.csv")
dfNa <- original
df <- na.omit(dfNa)
```

## What is mean total number of steps taken per day?

### Total steps per day
```{r}
total <- aggregate(df$steps, list(df$date), sum)
colnames(total) <- c('date', 'total_steps')
total
```

### Histogram
```{r fig.width=6, fig.height=4}
hist(total$total_steps, ylab = 'Number of Days', main="Total Steps Per Day")
```

### Mean and median
```{r}
mean(total$total_steps)
median(total$total_steps)
```

## What is the average daily activity pattern?
### Time series graph
```{r fig.width=6, fig.height=4}
daily <- aggregate(df$steps, list(df$interval), mean)
colnames(daily) <- c('interval', 'average')
plot(daily, type="l", main="Daily Activity Pattern", ylab = "Number of Steps", xlab = "5 min interval")
```

### Interval with the max value
```{r}
daily[daily$average == max(daily$average), ]
```

## Imputing missing values

### Total number of missing values
```{r}
sum(is.na(dfNa$steps))
```

### Fill the missing values with average interval
```{r fig.width=6, fig.height=4}
missingIntervals <- unique(dfNa[is.na(dfNa$steps), 'interval'])
for (i in 1:length(missingIntervals)) {
  interval <- missingIntervals[i]  
   dfNa[is.na(dfNa$steps) & dfNa$interval == interval, 'steps']  <- daily[daily$interval == interval, 'average']
}
total2 <- aggregate(dfNa$steps, list(dfNa$date), sum)
hist(total2$x, ylab = 'Number of Days', main="Total Steps Per Day (Imputed)")
```
### Mean and median
```{r}
mean(total2$x)
median(total2$x)
isSameMean <- mean(total2$x) == mean(total$total_steps)
isSameMedian <- median(total2$x) == median(total$total_steps)
isSameMean
isSameMedian
``` 

## Are there differences in activity patterns between weekdays and weekends?

### Add weekday/weekend column
```{r}
getDay <- function (d) {
  day <- as.character(weekdays(as.Date(d)))
  ifelse(day %in% c("Saturday", "Sunday"), 'weekend', 'weekday')
}
dfNa$day <- getDay(dfNa$date)
```

### Weekday-weekend plots
```{r fig.width=12, fig.height=8}
wd <- dfNa[dfNa$day == 'weekday', ]
we <- dfNa[dfNa$day == 'weekend', ]
wdDaily <- aggregate(wd$steps, list(wd$interval), mean)
weDaily <- aggregate(we$steps, list(we$interval), mean)

par(mfrow=c(2,1))
plot(wdDaily, type="l", main="Weekday", ylab = "Number of Steps", xlab = "5 min interval")
plot(weDaily, type="l", main="Weekend", ylab = "Number of Steps", xlab = "5 min interval")
```